\textsc{spiderman}: an open source code to model phase curves and secondary eclipses.

### 2016-10-13 ms.tex
written some methods

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
\begin{keywords}
planets and satellites: individual (WASP 52b)---stars: individual (WASP 52)---techniques: spectroscopic---planets and satellites: atmospheres---celestial mechanics---atmospheric effects
\end{keywords}

@@

General Circulation Models, GCM's, are currently the state of the art for predicting the wind patters and temperature distributions of the surface of exoplanets \citep[e.g.]{Showman2008}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{img/zhang_quad.pdf}
\caption{This is a caption}
\label{fig:zhang_quad}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{img/zhang_lc.pdf}
\caption{This is a caption}
\label{fig:extract_region}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{img/zhang_lc_zoom.pdf}
\caption{This is a caption}
\label{fig:extract_region}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{img/free_parameterstriangle.pdf}
\caption{This is a caption}
\label{fig:extract_region}
\end{center}
\end{figure}

\begin{figure*}
\begin{center}
\includegraphics[width=\columnwidth]{img/free_parametersflux_map.pdf}
\includegraphics[width=\columnwidth]{img/free_parametersflux_errs.pdf}
\caption{left: right: the constraints on the brightness distribution.}
\label{fig:extract_region}
\end{center}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=\columnwidth]{img/free_parameterstemp_map.pdf}
\includegraphics[width=\columnwidth]{img/free_parameterstemp_errs.pdf}
\caption{The underlying temperature distributions generated by the zhang model.}
\label{fig:extract_region}
\end{center}
\end{figure*}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{img/precision.pdf}
\caption{This is a caption}
\label{fig:extract_region}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{img/exec_time.pdf}
\caption{This is a caption}
\label{fig:exec_time}
\end{center}
\end{figure}

@@

Some physical models, such as the one adapted from \citep{zhang2016} are defined in terms of temperature, where of course the observable that is being fit for is the emitted flux. Therefore, it is necessary

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{img/system.pdf}
\caption{This is a caption}
\label{fig:extract_region}
\end{center}
\end{figure}

@@

As a test of the performance of the code the model was applied to a phase curve and secondary eclipse of WASP-43b taken with HST WFC3. These data had previously been published in \citet{Stevenson2014c}. However, in this paper the phase curve and the secondary eclipse are treated as seperate phenomena, when of course, they are both manifestations of the same underlying brightness distribution. Indeed, the bright dayside of the planet is the main driver of the phase curve, and is also the part occulted by the stellar disc. It is therefore more parsimonious, if at all possible, to use a model that accounts for both of these observations simultaneously. This also reduces the total number of paramters needed, which is satisfying from an information content angle.

@@

WASP-43b appears to have a near isothermal profile \citep{Stevenson2014c}, which should correspond to a zero limb darkening. However, as I show in section X. there is a preference in the fit for a \emph{limb brightened profile} which may indicate the presence of a weak thermal inversion layer, or may simply be due to oversimplifications in the model fit.

@@

The systematics model is the same that was used in the previous treatment \citep{Stevenson2014c}.

@@

There is an offset in the recorded flux depending on which way the the ccd was scanning during the readout, in one half of exposures the scan direction is in the same direction as the readout, and in the other half it is the offset. This leads to a very repeatable offset in the fluxes between the two scan directions, which can be captured with a single parameter "scale"

@@

@@

Each visit displays a decreasing slope, which is modelled as a second order polynomial. The cause of these features is unknown, and they have not been linked to any physical parameters of the telescope \cite{Wakeford2016}. Here, they are also likely blended with the "Breathing" systematics of the instrument, which are caused by the changing thermal environment of the telescope during its orbit changing the focal length of the instrument. The shape of these visit long effects is slightly different between the three visits used here, so the three paramters of the polynomial are fit seperately for the three visits.

@@

@@

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{img/huge_triangle.pdf}
\caption{stupidly big triangle plot1}
\label{fig:extract_region}
\end{center}
\end{figure*}

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{img/new_lc.pdf}
\caption{The final fitted lightcurve. The primary transit is fit by \textsc{batman} and is off the scale of this figure.}
\label{fig:extract_region}
\end{center}
\end{figure}

@@

- Confirming the detected flux on the nightside is a real feature and not an artefact
- Further test for degeneracies between HST systematics and the physical model of the phase curve, in particular the combination of visit long slope and breathing modes. This could be accomplished by a Gaussian process analysis.

>>>>

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
\begin{keywords}
%planets and satellites: individual (WASP 43b)---stars: individual (WASP 43)---techniques: spectroscopic---planets and satellites: atmospheres---celestial mechanics---atmospheric effects
\end{keywords}

@@

The planets are hot enough that they have appreciable infrared flux emission, with contrasts with their parents stars of order 0.1\%, making detection with current generation instruments feasible. \citet{Deming2005} detected the infrared flux of HD\,209458b by observing the secondary transit of the planet with the MIPS instrument on \emph{Spitzer}. Using the IRAC instrument on \emph{Spitzer}, \citet{Knutson2007b} went on to measure the secondary eclipse of HD\,189733b with high precision, and also show that the system displayed a sinusoidal \emph{phase curve}, caused by the bright regions of the planet rotating into and out of view. Intriguingly, the phase of this sinusoidal curve was significantly offset from the time of secondary eclipse, validating earlier predictions from GCM's.

Excitingly, it is possible to invert phase curves to give information on the longitudinal brightness distribution of the exoplanet, providing a way to map the surfaces of other worlds.

\cite{Majeau2012} showed that as well as mapping longitudinally with phase curves, it is possible to get a two dimensional map of the dayside of an exoplanet from the secondary eclipse, provided that there is a non zero impact parameter. This technique works because the inclination of the system results in the occulting limb of the star scanning across the disc of the planet in both longitude and latitude during ingress and egress. The brightness distribution of the planet's permanent dayside will be imprinted on the precise shape of the ingress and egress curves.

There appear to be, broadly, two classes of exoplanet atmospheres with regards to heat transport. $\mu$ Andromedae, HD\,179949b and HD\,209458b are all significantly hotter than HD\,189733b and display much higher temperature contrasts between the day and night sides, implying inefficient redistribution of energy. \citep{Harrington2006,Cowan2007,zellem2014}. WASP-43b is similarly an extremely hot Jupiter, so would be expected to have high temperature contrast between day and night sides. The primary and secondary eclipses of this planet, as well as the phase curve, have been measured with HST WFC3 and reported in \citep{Stevenson2014}, and these data will be reanalysed here.

The high velocity winds transporting this energy have been predicted to be on the order of kilometers per second, which would leave a measurable imprint on the spectral absorption lines in the planets atmosphere through Doppler shifting \citep[e.g.][]{Showman2013}, and indeed, this effect has been observed in the transit of HD\,209458b \citep{Snellen2010} and HD\,189733b \citep{Louden2015}.

\citep[e.g.][]{Showman2015}

General Circulation Models (GCM's), are currently the state of the art for predicting the wind patters and temperature distributions of the surface of exoplanets \citep[e.g.][]{Showman2008}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{img/zhang_quad.pdf}
\caption{An example of the temperature distribution generated by using the Zhang model. Clockwise from top left: Phase = 0.0,0.25,0.75,1.0}
\label{fig:zhang_quad}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{img/zhang_lc.pdf}
\caption{An example lightcurve generated using the distribution in figure \ref{fig:zhang_quad}}
\label{fig:zhang_lc}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{img/zhang_lc_zoom.pdf}
\caption{A close up of the secondary eclipse in figure \ref{fig:zhang_lc}. Note the different shape of the ingress and egress, caused by the asymmetry in the flux distribution on the planet.}
\label{fig:zhang_zoom}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{img/free_parameterstriangle.pdf}
\caption{A section of the larger corner plot in figure \ref{fig:big triangle}. The parameters for the Zhang model fit are loosely correlated. The distribution for the stellar temperature, $T_s$, is set by the prior}
\label{fig:triangle corner sub}
\end{center}
\end{figure}

\begin{figure*}
\begin{center}
\includegraphics[width=\columnwidth]{img/free_parametersflux_map.pdf}
\includegraphics[width=\columnwidth]{img/free_parametersflux_errs.pdf}
\caption{left: right: the constraints on the brightness distribution.}
\label{fig:best_fit_flux}
\end{center}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=\columnwidth]{img/free_parameterstemp_map.pdf}
\includegraphics[width=\columnwidth]{img/free_parameterstemp_errs.pdf}
\caption{The underlying temperature distributions generated by the zhang model.}
\label{fig:best_fit_temp}
\end{center}
\end{figure*}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{img/precision.pdf}
\caption{A comparison between a uniform flux distribution eclipse model generated with 400 segments and a single segment. If the integrator is correct, then there should be no difference between these two models. The bottom panel shows that the results are identical, apart from errors at the level of floating point precision (dotted lines) during ingress and egress.}
\label{fig:precision}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{img/exec_time.pdf}
\caption{The speed with which a model with 1000 timepoints can be generated}
\label{fig:exec_time}
\end{center}
\end{figure}

@@

Some physical models, such as the one adapted from \citep{Zhang2016} are defined in terms of temperature, where of course the observable that is being fit for is the emitted flux. Therefore, it is necessary

@@

typing the command

\begin{Verbatim}[frame=single]
> spider_params.plot\_system()
\end{Verbatim}

will generate a plot like figure \ref{fig:plot_system}

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{img/system.pdf}
\caption{An example of the graphical output that \textsc{spiderman} can produce. This image is a scale representation of the system, with the planetary flux map included. This is useful for visualising the model. It was generated with the spiderman.plot\_system() command. The image of the star is purely aesthetic, and does not contain any model detail.}
\label{fig:plot_system}
\end{center}
\end{figure}

\textsc{spiderman} has been designed to ensure that it is easy to visualise the appearance of the model at all stages.

@@

As a test of the performance of the code the model was applied to a phase curve and secondary eclipse of WASP-43b taken with HST WFC3. These data had previously been published in \citet{Stevenson2014}. However, in this paper the phase curve and the secondary eclipse are treated as seperate phenomena, when of course, they are both manifestations of the same underlying brightness distribution. Indeed, the bright dayside of the planet is the main driver of the phase curve, and is also the part occulted by the stellar disc. It is therefore more parsimonious, if at all possible, to use a model that accounts for both of these observations simultaneously. This also reduces the total number of paramters needed, which is satisfying from an information content angle.

@@

WASP-43b appears to have a near isothermal profile \citep{Stevenson2014}, which should correspond to a zero limb darkening. However, as I show in section X. there is a preference in the fit for a \emph{limb brightened profile} which may indicate the presence of a weak thermal inversion layer, or may simply be due to oversimplifications in the model fit.

@@

The systematics have effects on scales of \emph{visits}, \emph{orbits} and individual exposures, where a visit is a continuous set of orbits directed at a target. Due to the low earth orbit of Hubble, each complete orbit lasts 95 minutes, and for approximately half of this time the target will be unobservable due to being behind the Earth. For this reason, a complete exoplanet transit can not be observed with one "visit", and it is necessary to observe several transits with different temporal phasing to completely cover the full orbital phase of the system.

In this study, three visits, each with 12 orbits was used to cover the full phase of WASP-43. The phasing can be seen in figure \ref{fig:systematics}, and a phase folded lightcurve can be seen in figure \ref{fig:phase folded}

The systematics model we use to fit WASP-43 is the same that was used in the previous treatment \citep{Stevenson2014}.

@@

There is an offset in the recorded flux depending on which way the the ccd was scanning during the readout, in one half of exposures the scan direction is in the same direction as the readout, and in the other half it is the offset. This leads to a very repeatable offset in the fluxes between the two scan directions, which can be captured with a single parameter \emph{scale}

@@

Each orbit displays a characteristic exoplential "hook" signal, which can be modelled as

@@

Each \emph{visit} displays a decreasing slope, which is modelled as a second order polynomial. The cause of these features is unknown, and they have not been linked to any physical parameters of the telescope \cite{Wakeford2016}. Here, they are also likely blended with the "Breathing" systematics of the instrument, which are caused by the changing thermal environment of the telescope during its orbit changing the focal length of the instrument. The shape of these visit long effects is slightly different between the three visits used here, so the three paramters of the polynomial are fit seperately for the three visits.

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{img/new_lc.pdf}
\caption{The final fitted lightcurve. The primary transit is fit by \textsc{batman} and is off the scale of this figure.}
\label{fig:phase folded}
\end{center}
\end{figure}

@@

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{img/huge_triangle.pdf}
\caption{stupidly big triangle plot1}
\label{fig:big triangle}
\end{center}
\end{figure*}

@@

- Confirming the detected flux on the nightside is a real feature and not an artefact
- Further test for degeneracies between HST systematics and the physical model of the phase curve, in particular the combination of visit long slope and breathing modes. This could be accomplished by a Gaussian process analysis.
- Perform recovery with a non-informative model, such as spherical harmonics, and see if the recovered brightness or temperature maps are the same as recovered through physical models.
- Investigating the apparent preference for limb brightened profiles in the model fit. This may be evidence of
